On the formation of H$\alpha$ around classical T Tauri stars

### 2006-01-03 TTau_Halpha/kurosawa.tex
Fixed typos and grammatical errors.  (Ryuichi)
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@609 25464f21-6099-402c-b30a-640f1d40d38b

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,useAMS, usenatbib]{mn2e}
%\usepackage[T1]{fontenc}
%\usepackage[latin1]{inputenc}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amssymb}

@@

\usepackage{babel}
\makeatother
\begin{document}
\title[On H-$\alpha$ formation of T Tauri stars]{On formation of H-$\alpha$ from classical T Tauri stars: the disc, wind, and magnetospheric-accretion hybrid model}

\author[R. Kurosawa et\,al.]{Ryuichi Kurosawa\thanks{E-mail: rk@astro.ex.ac.uk}, Tim J. Harries and Neil H. Symington\\ School of Physics, University of Exeter, Stocker Road, Exeter EX4 4QL}

@@

We present disc--wind--magnetosphere hybrid radiative transfer models
for classical T~Tauri stars, and investigate the formation of H$\alpha$
to understand their complex circumstellar environment, and to understand
the wide variety of emission line profiles seen in observations. To
overcome the limited applicability of the radiative transfer models
with only magnetospherical accretion to fitting observed H$\alpha$
profiles, and to any other lines affected by the stellar wind) presented,
two types of kinematic wind models are introduced: (1)~the bipolar
wind model in which the wind is originated from the star itself, and
(2)~the disc-wind model in which the wind originates from the inner
part of the accretion disc. We perform systematic investigations of
the model parameters for the wind and the magnetosphere to search
for possible geometrical and physical conditions which lead to the
types of the profiles seen in observations. We find both wind models
can reproduce the wide range profile types seen is observations. We
also find the inclination dependency of the line equivalent width
predicted by the bipolar wind model agree with trends seen in the
observation, but the disc-wind model does not. Using the model results,
we examine the H$\alpha$ spectroscopic classification used by Reipurth
et.~al , and discuss the basic physical conditions required to reproduce
the profiles in each classified type. Using the hybrid model, we fit
the observed H$\alpha$ from the optical component of T~Tau triple
system (T~Tau~N), and find its ratio of the mass-loss to mass-accretion
rate to be $\sim0.18$ with the cold ($T_{\mathrm{wind}}\approx8000\,\mathrm{K}$)
bipolar wind model .

@@

T~Tauri stars (TTS) are young ($<\sim3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass stars, and known as progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some
of the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit an excess
amount of continuum flux in the ultraviolet (UV) and infrared (IR).
Their spectral energy distribution and polarisation data suggest the
presence of a circumstellar disc, and it plays an important role in
regulating dynamics of gas flows around CTTS.

Many observational studies (e.g., \citealt{herbig:1962}; \citealt{edwards:1994};
\citealt{kenyon:1994}; \citealt{reipurth:1996}; \citealt{alencar:2000})
of CTTS line profiles show evidences for both outward wind flows and
inward accretion flows, as seen in the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC) profiles
respectively. Typical mass-loss rates of CTTS are about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$
to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ (e.g., \citealt{kuhi:1964};
\citealt{edwards:1987}; \citealt{hartigan:1995}), and the mass-accretion
rates are also about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$
(e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988}; \citealt{gullbring:1998}).
Recent H$\alpha$ spectro-astrometric observations by \citet{takami:2003}
show the direct evidence for the presence of bipolar and monopolar
outflows down to $\sim1\,\mathrm{AU}$ scale (e.g.\,CS~Cha and RU~Lup).
Similarly, ESO VLT observation of high-resolution ($\mathrm{R=50\,000}$)
two-dimensional spectral of edge-on CTTS (HH30$^{*}$, HK~Tau~B,
and HV~Tau~C) by \citet{appenzeller:2005} show the extended H$\alpha$
emission in the direction perpendicular to the obscuring circumstellar
disc, and in both above and below the disc --- suggesting the presence
of the bipolar outflows. In slightly larger scale, \emph{HST} observation
of HH30 \citet{burrows:1996} shows the jet traced to within $\lesssim30$~AU
from the star. The jet has a cone shape with an opening angle of $3^{\circ}$
between 70 and 700 AU \citep{koenigl:2000}. \citet{alencar:2000}
found about 80 per cent of their samples (30 CTTS) show blue-shifted
absorption components in at least one of Balmer lines and Ca~K (most
commonly in H$\alpha$).

In a currently favoured model of accretion flows around CTTS, the
accretion discs are disrupted by the magnetosphere of stars which
channels the gas from the disc onto the surface of the stars (e.g.,
\citealt{uchida:1985}; \citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture of the accretion flows is supported
by the evidences that CTTS have relatively strong ($\sim10^{3}\mathrm{G}$)
magnetic field (e.g., \citealt*{jonhs-krull:1999}; \citealt{guenther:1996})
and by the radiative transfer models which reproduce the observed
profiles for some TTS (\citealt{muzerolle:2001}). The magnetospherical
accretion model naturally explains the blue-ward asymmetric emission
lines (seen in some of CTTS) caused by the partial occultation of
the flow by the disc, and the redshifted absorption component at the
typical free-fall velocities (a few hundred $\mathrm{km\, s^{-1}}$)
seen in some of CTTS. Despite the success of the magnetospherical
accretion model in explaining the line profiles in some CTTS, the
overwhelming observational evidences for the outflow (mentioned above)
in the CTTS profiles suggests that this model is only a part of a
whole picture. Clearly, the modification to include the out-flowing
wind/jet flow is necessary if one requires to predict the mass-accretion
rate and the mass-loss rate of CTTS by modelling their emission profiles
(e.g. H$\alpha$).

@@

The main aim of this paper is to find a simple kinematic model which
can reproduce the wide variety of the observed profiles, and to perform
the empirical studies of the line formation. More specifically, we
present disc--wind--magnetosphere hybrid models for CTTS, and perform
parametric studies of the H$\alpha$ formation. This study should
provide preliminary physical conditions which lead to the wide variety
of emission line profiles seen in the observation, and would help
to construct more comprehensive circumstellar models (e.g. via MHD
simulations) of T~Tauri stars. the Using the model results, we examine
the H$\alpha$ spectroscopic classification used by \citet{reipurth:1996},
and compare models with observations. We will also discuss whether
our model is consistent with some predictions made by the recent MHD
studies i.e.~ $\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$
(e.g.~\citealt{koenigl:2000}).

@@

To understand how the different part of the CTTS circumstellar environment
contributes to the formation of H$\alpha$, the model spaced is divided
into four different regions: 1.~Central continuum source, 2.~magnetosphere
accretion flow, 3.~wind outflow, and 4.~accretion disc. Fig.~\ref{fig:config_dipolar}
depicts the relative location in the model space. In all regions,
the density is symmetric around the z-axis. The innermost radius of
the magnetosphere at the equatorial plane coincides with the the inner
radius of the accretion disc. From the inner most part of the accretion
disc, the gas freely fall, moving along the magnetic filed onto the
surface of the star. For the purpose of computational simplicity,
the collimated bipolar wind are defined only in the outside of the
largest radius of the magnetosphere (to avoid overlapping of the wind
with the magnetosphere). In the following subsections, the details
of model components will be described. In our models, two different
types of the wind configurations are considered: (1)~the bipolar
wind resembling the density structure of the MHD simulation by \citealt{krasnopolsky:2003}
in large scale ($\sim100$~AU) in section~\ref{sub:wind_jet}, and
(2) that in small scale ($\sim10$~AU) in section~\ref{sub:Alternative-model-disc-wind}.

@@

\caption{Basic Model Configuration. The system consist of four components: (1)~the continuum source located at the origin $\left( o \right)$ of the cartisian coordinates $\left(x,y,z\right)$ -- the y-axis are into the paper, (2)~magnetospherical accretion flow, (3)~bipolar wind outflow, and (4)~accretion disc. The density distribution is symmetric around the z-axis. The innermost raius of the magnetsphere (at the equotrial plane) coincides with the the inner radius of the accretion disc. From the inner most part of the accretion disc, the gas freely fall, moving along the magneic filed onto the surface of the star. For a computational simplicity, the bipolar wind starts from just outside of the largest radius ($R_{\mathrm{mo}}$) of the magnesophere (dotted line). }

@@

For a model which includes magnetospherical accretion, an additional
source of the continuum source is considered. As the in-falling gas
becomes closer to the stellar surface, it is decelerates in a strong
shock, and heated to $\sim10^{6}\,\mathrm{K}$ (with typical parameters).
The X-ray radiation produced in the shock will be absorbed by the
gas located near, and re-emitted in optical and UV radiation (\citealt{koenigl:1991};
\citealt{hartmann:1994}). This will create hot rings near the stellar
surface where the magnetic field intersects with the surface. For
simplicity, the free-falling kinetic energy is assumed to be thermalized
in the radiating layer, and re-emits as blackbody radiation (with
a single temperature). With the parameters of the magnetosphere and
the star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings, and its temperature is
about $6400\,\mathrm{K}$. If the mass-accretion rate is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$,
the ratio of this accretion luminosity to the photospheric luminosity
is about $0.5$. The continuum emission from the hot rings is taken
into account when computing the line profiles.

@@

The magnetic field and the gas stream lines are assumed to have the
following simple form: \begin{equation}
r=R_{\mathrm{m}}\sin^{2}\theta\label{eq:dipole}\end{equation}
 \citep[see][]{ghosh:1977} where $r$, and $\theta$ are coordinates
of the field point ($p$) in Fig.~\ref{fig:config_dipolar} in spherical
coordinates, and $R_{\mathrm{m}}$is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$). The range of $R_{\mathrm{m}}$is
restricted between $R_{\mathrm{mi}}$ and $R_{\mathrm{mo}}$. Using
the field geometry above and the conservation of energy, the velocity
and the density of the accreting gas along the steam line are found
as in \citet{hartmann:1994}.

@@

As mentioned earlier, the wind model presented here is to simulate
the collimated MHD disc-wind model of \citet{krasnopolsky:2003} in
large scale ($\sim100$~AU) , and the alternative wind model which
resembles the model of \citet{krasnopolsky:2003} in small scale ($\sim10$~AU)
will be presented later (section~\ref{sub:Alternative-model-disc-wind}).

Similar to the simple wind model of \citet{appenzeller:2005}, the
following parametrisation of collimated bipolar wind (region 3 in
Fig.~\ref{fig:config_dipolar}) is adopted. The wind velocity field,
$\mathbf{v}_{\mathrm{wind}}\left(r,\theta\right)$, is consist of
radial and azimuthal components which depend on the spherical coordinates
$r$ and $\theta$. The radial component $v_{r}\left(r\right)$ is
assumed to be in the classical beta-velocity law \citep[c.f.][]{castor:1979},
and the azimuthal component $v_{\phi}$ is assumed to be a constant
fraction ($\gamma$) of the Keplerian velocity for a given distance
($w=\sqrt{x^{2}+y^{2}}$) from the symmetry axis ($z$-axis) , i.e.:

\begin{eqnarray}
\mathbf{v}_{\mathrm{wind}} & = & v_{r}\hat{\mathbf{r}}+v_{\phi}\hat{\mathbf{\bphi}}\label{eq:wind_velocity_def}\end{eqnarray}
where \begin{equation}
v_{r}\left(r\right)=v_{r0}+\left(v_{\infty}-v_{r0}\right)\left(1-\frac{R_{\mathrm{mo}}}{r}\right)^{\beta}\,,\label{eq:wind_velocity_radial_part}\end{equation}
\begin{equation}
v_{\phi}\left(w\right)=\gamma\,\left(\frac{GM_{*}}{w}\right)^{1/2}\,,\label{eq:wind_velocity_azimuth_part}\end{equation}
Note that the base of the wind starts at $r=R_{\mathrm{mo}}$. This
is chosen so mainly for computational simplicity i.e. to avoid the
overlap between the wind and the magnetosphere; however, see \citet{matt:2005}
for the possibility that wind originates from the stellar surface
(restricted near the polar caps). The range of the polar angle for
the wind is restricted to $\theta>\left|\theta_{\mathrm{disc}}\right|$
where $\theta_{\mathrm{disc}}$ is the opening angle of the accretion
disc, to avoid the overlap. $v_{r0}$is the small radial velocity
at the base of the wind ($r=R_{\mathrm{mo}}$). Normally, $v_{r0}=10\,\mathrm{km\, s^{-1}}$
which is approximately equal to the thermal velocity of hydrogen with
$T=7500\,\mathrm{K}$. Following \citet{appenzeller:2005}, $\gamma=0.05$
is adopted. The dependency of $v_{r}$ in polar direction ($\theta=0$)
on the values of wind acceleration parameter $\beta$ is shown in
Figure~\ref{fig:wind_vr_rho}. All other parameters describing the
wind are fixed as the standard values given in Table~\ref{tab:std_parameters}.

@@

\caption{The dependency of the bipolar wind velocity and density structures on the wind accerelation parameter $\beta$. The radial component of the wind  velocity (equation~\ref{eq:wind_velocity_radial_part}) in polar direction, $\theta = 0$ as a function of radius is shown on left. The wind density (equation~\ref{eq:wind_density_final} along polar direction as a function of radius is shown on right. For all $\beta$ values, the density is normalized to the desnity at the base of the wind in polar direction, i.e. $\rho_0 = \rho\left(R_\mathrm{mo},0\right)$.  The smaller the value of $\beta$, the faster the accerelation of the wind. For the raidus ($r/R_{\mathrm{mo}} < 10$, both velocity and density are sensitive to the value of $\beta$. For all $\beta$ values, the radial velocity of the wind reaches the terminal velocity ($\sim 210\,\kmps$ by  $r/R_{\mathrm{mo}} \sim 1000$. The initial velocity $V_{0}=10\,\kmps$, which approximately corresponds to the thermal velocity of a hydrogen atom at 7500~K,  is used for all $v_{r}$ plots. Beyound $r/R_{\mathrm{mo}}= 100$, little difference is seen in the polar density with diffrent $\beta$ values.}

@@

To calculate the dust scattering and absorption cross section as a
function of wavelength, the optical constants of \citet{draine:1984}
for amorphous carbon grains and \citet{hanner:1988} for silicate
grains are used. The model uses the {}``large grain'' dust model
of \citet{wood:2002} in which the dust grain size distribution is
described by the following function:\begin{equation}
n\left(a\right)da=\left(C_{\mathrm{C}}+C_{\mathrm{Si}}\right)\, a^{-p}\exp\left[-\left(\frac{a}{a_{c}}\right)^{q}\right]da\label{eq:grain-dist-function}\end{equation}
 where $a$ is the grain size restricted between $a_{\mathrm{min}}$
and $a_{\mathrm{max}}$, and $C_{\mathrm{C}}$ and $C_{\mathrm{Si}}$
are the terms set by requiring the grains to completely deplete a
solar abundance carbon and silicon. The parameters adopted in our
model are: $C_{\mathrm{C}}=1.32\times10^{-17}$, $C_{\mathrm{Si}}=1.05\times10^{-17}$,
$p=3.0$, $q=0.6$, $a_{\mathrm{\mathrm{min}}}=0.1\,\mathrm{\mathrm{\mathrm{\mu m}}}$,
$a_{\mathrm{\mathrm{max}}}=1000\,\mathrm{\mathrm{\mathrm{\mu m}}}$,
and $a_{c}=50\,\mathrm{\mathrm{\mathrm{\mu m}}}$. This corresponds
to Model~1 of the dust model used by \citet{wood:2002}. See also
their Fig.~3 The relative number of each grain is assumed to be that
of solar abundance, C/H$\sim3.5\times10^{-4}$ \citep{anders:1989}
and Si/H$\sim3.6\times10^{-5}$ \citep{grevesse:1993} which are similar
to values found in the ISM model of \citet{mathis:1977} and \citet*{kim:1994}.
Similar abundances were used in the circumstellar disc models of \citet{cotera:2001}.

@@

\caption{Basic Model Configuration of the disc-wind, magnetosphere hybrid model. The system consist of four components: (1)~the continuum source located at the origin $\left( o \right)$ of the cartisian coordinates $\left(x,y,z\right)$ -- the y-axis are into the paper, (2)~magnetospherical accretion flow, (3)~colliminated wind/jet outflow, and (4)~geometrically thin (but opaque) accretion disc. The disc wind originates from the disc surface between $w_{i}=R_{wi}$ and $w_{i}=R_{wo}$.  The wind source points ($S$), from which the stream lines diverges, are placed at distance $d$ above and below the star. By changing the value of $d$,  the degree of the wind collimation is controlled. The density distribution is symmetric around the z-axis. }

>>>>

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amssymb}

@@

\usepackage{babel}
\makeatother
\begin{document}
\title[On the formation of H$\alpha$ around classical T Tauri stars]{On the formation of H$\alpha$ around classical T Tauri stars}

\author[R. Kurosawa et\,al.]{Ryuichi Kurosawa$^1$\thanks{E-mail: rk@astro.ex.ac.uk}, Tim J. Harries$^1$ and Neil H. Symington$^2$\\
$^1$School of Physics, University of Exeter, Stocker Road, Exeter EX4 4QL. \\
$^2$School of Physics and Astronomy, University of St. Andrews, North Haugh, St. Andrews, Fife, KY16 9SS.}

@@

We present  radiative transfer models of classical T~Tauri stars, and investigate the formation of H$\alpha$
in order to understand their complex circumstellar environment, and to explain
the wide variety of emission line profiles seen in observations. To
overcome the limited applicability of the radiative transfer models
with magnetospherical accretion alone to fitting observed H$\alpha$
profiles, and to any other lines affected by the stellar wind) presented,
two types of kinematic wind models are introduced: (1)~the bipolar
wind model in which the wind is originated from the star itself, and
(2)~the disc-wind model in which the wind originates from the inner
part of the accretion disc. We perform systematic investigations of
the model parameters for the wind and the magnetosphere to search
for possible geometrical and physical conditions which lead to the
types of profiles seen in observations. We find both wind models
can reproduce the wide range profile types seen is observations, 
although the inclination dependence of the line equivalent width
predicted by the bipolar wind model agree with trends seen in the
observation, but the disc-wind model does not. Using the model results,
we examine the H$\alpha$ spectroscopic classification used by Reipurth
et.~al , and discuss the basic physical conditions required to reproduce
the profiles in each classified type. Using the hybrid model, we fit
the observed H$\alpha$ from the optical component of T~Tau triple
system (T~Tau~N), and find its ratio of the mass-loss to mass-accretion
rate to be $\sim0.18$ with the cold ($T_{\mathrm{wind}}\approx8000\,\mathrm{K}$)
bipolar wind model .

@@

T~Tauri stars (TTS) are young ($<\sim 3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass stars, and are the progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some
of the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit an excess
amount of continuum flux in the ultraviolet (UV) and infrared (IR).
Their spectral energy distribution and polarisation data suggest the
presence of a circumstellar disc, which plays an important role in
regulating dynamics of gas flows around CTTS.

Many observational studies (e.g., \citealt{herbig:1962}; \citealt{edwards:1994};
\citealt{kenyon:1994}; \citealt{reipurth:1996}; \citealt{alencar:2000})
of CTTS line profiles show evidence for both outward wind flows and
inward accretion flows, as seen in the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC) profiles
respectively. Typical mass-loss rates of CTTS are about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$
to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ (e.g., \citealt{kuhi:1964};
\citealt{edwards:1987}; \citealt{hartigan:1995}), and the mass-accretion
rates are also about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$
(e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988}; \citealt{gullbring:1998}).
Recent H$\alpha$ spectro-astrometric observations by \citet{takami:2003}
show the direct evidence for the presence of bipolar and monopolar
outflows down to $\sim 1\,\mathrm{AU}$ scale (e.g.\,CS~Cha and RU~Lup).
Similarly, ESO VLT observation of high-resolution ($R=50\,000$)
two-dimensional spectra of edge-on CTTS (HH30$^{*}$, HK~Tau~B,
and HV~Tau~C) by \citet{appenzeller:2005} show the extended H$\alpha$
emission in the direction perpendicular to the obscuring circumstellar
disc, and in both above and below the disc --- suggesting the presence
of the bipolar outflows. On a larger scale, \emph{HST} observation
of HH30 \citep{burrows:1996} shows the jet traced to within $\lesssim30$~AU
from the star. The jet has a cone shape with an opening angle of $3^{\circ}$
between 70 and 700 AU \citep{koenigl:2000}. \citet{alencar:2000}
found about 80 per cent of their samples (30 CTTS) show blue-shifted
absorption components in at least one of Balmer lines and Ca~K (most
commonly in H$\alpha$).

In the currently favoured model of accretion flows around CTTS, the
accretion discs are disrupted by the magnetosphere of stars which
channels the gas from the disc onto the surface of the stars (e.g.,
\citealt{uchida:1985}; \citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture of the accretion flows is supported
by the evidences that CTTS have relatively strong ($\sim 10^{3}\mathrm{G}$)
magnetic field (e.g., \citealt*{jonhs-krull:1999}; \citealt{guenther:1996})
and by the radiative transfer models which reproduce the observed
profiles for some TTS (\citealt{muzerolle:2001}). The magnetospherical
accretion model naturally explains the blue-ward asymmetric emission
lines (seen in some of CTTS) caused by the partial occultation of
the flow by the disc, and the redshifted absorption component at the
typical free-fall velocities (a few hundred $\mathrm{km\, s^{-1}}$)
seen in some of CTTS. Despite the success of the magnetospherical
accretion model in explaining the line profiles in some CTTS, the
overwhelming observational evidences for the outflow (mentioned above)
in the CTTS profiles suggests that this model is only a part of a
complete picture. Clearly, the modification to include the out-flowing
wind/jet flow is necessary if one requires to predict the mass-accretion
rate and the mass-loss rate of CTTS by modelling their emission profiles
(e.g. H$\alpha$).

@@

The main aim of this paper is to find a simple kinematic model which
can reproduce the wide variety of the observed profiles, and to perform
the empirical studies of the line formation. More specifically, we
present disc--wind--magnetosphere hybrid models for CTTS, and perform
parametric studies of the H$\alpha$ formation. This study should
provide preliminary physical conditions which lead to the wide variety
of emission line profiles seen in the observation, and would help
to construct more comprehensive circumstellar models (e.g. via MHD
simulations) of T~Tauri stars. Using the model results, we examine
the H$\alpha$ spectroscopic classification used by \citet{reipurth:1996},
and compare models with observations. We will also discuss whether
our model is consistent with some predictions made by the recent MHD
studies i.e.~ $\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$
(e.g.~\citealt{koenigl:2000}).

@@

To understand how the different part of the CTTS circumstellar environment
contributes to the formation of H$\alpha$, the model space is divided
into four different regions: (1)~a central continuum source, (2)~magnetosphere
accretion flow, (3)~wind outflow, and (4)~accretion disc. Fig.~\ref{fig:config_dipolar}
depicts the relative location in the model space. In all regions,
the density is symmetric around the z-axis. The innermost radius of
the magnetosphere at the equatorial plane coincides with the the inner
radius of the accretion disc. From the inner most part of the accretion
disc, the gas freely fall, moving along the magnetic filed onto the
surface of the star. For the purpose of computational simplicity,
the collimated bipolar wind are defined only in the outside of the
largest radius of the magnetosphere (to avoid overlapping of the wind
with the magnetosphere). In the following subsections, the details
of model components will be described. In our models, two different
types of the wind configurations are considered: (1)~the bipolar
wind resembling the density structure of the MHD simulation by \citealt{krasnopolsky:2003}
in large scale ($\sim 100$~AU) in section~\ref{sub:wind_jet}, and
(2) that in small scale ($\sim 10$~AU) in section~\ref{sub:Alternative-model-disc-wind}.

@@

\caption{Basic Model Configuration. The system consist of four
components: (1)~the continuum source located at the origin $\left( o
\right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
y-axis are into the paper, (2)~magnetospherical accretion flow,
(3)~bipolar wind outflow, and (4)~accretion disc. The density
distribution is symmetric around the z-axis. The innermost radius of
the magnetosphere (at the equatorial plane) coincides with the the
inner radius of the accretion disc. From the inner most part of the
accretion disc, the gas freely fall, moving along the magnetic filed
onto the surface of the star. For a computational simplicity, the
bipolar wind starts from just outside of the largest radius
($R_{\mathrm{mo}}$) of the magnetosphere (dotted line). }

@@

For a model which includes magnetospherical accretion, an additional
source of the continuum source is considered. As the in-falling gas
becomes closer to the stellar surface, it is decelerates in a strong
shock, and heated to $\sim 10^{6}\,\mathrm{K}$ (with typical parameters).
The X-ray radiation produced in the shock will be absorbed by the
gas located near, and re-emitted in optical and UV radiation (\citealt{koenigl:1991};
\citealt{hartmann:1994}). This will create hot rings near the stellar
surface where the magnetic field intersects with the surface. For
simplicity, the free-falling kinetic energy is assumed to be thermalized
in the radiating layer, and re-emits as blackbody radiation (with
a single temperature). With the parameters of the magnetosphere and
the star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings. If the mass-accretion rate is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$,
the ratio of this accretion luminosity to the photospheric luminosity
is about $0.5$, and the corresponding temperature of the hot rings is
about $6400\,\mathrm{K}$.  The continuum emission from the hot rings is taken
into account when computing the line profiles.

@@

The magnetic field and the gas stream lines are assumed to have the
following simple form: \begin{equation}
r=R_{\mathrm{m}}\sin^{2}\theta\label{eq:dipole}\end{equation}
 \citep[see][]{ghosh:1977} where $r$, and $\theta$ are coordinates
of the field point ($p$) in Fig.~\ref{fig:config_dipolar} in spherical
coordinates, and $R_{\mathrm{m}}$is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$). The range of
$R_{\mathrm{m}}$ is
restricted between $R_{\mathrm{mi}}$ and $R_{\mathrm{mo}}$. Using
the field geometry above and the conservation of energy, the velocity
and the density of the accreting gas along the steam line are found
as in \citet{hartmann:1994}.

@@

As mentioned earlier, the wind model presented here is to simulate
the collimated MHD disc-wind model of \citet{krasnopolsky:2003} in
large scale ($\sim 100$~AU) , and the alternative wind model which
resembles the model of \citet{krasnopolsky:2003} in small scale ($\sim
10$~AU)
will be presented later (section~\ref{sub:Alternative-model-disc-wind}).

Similar to the simple wind model of \citet{appenzeller:2005}, the
following parametrisation of collimated bipolar wind (region 3 in
Fig.~\ref{fig:config_dipolar}) is adopted. The wind velocity field,
$\boldsymbol{v}_{\mathrm{wind}}$, is consist of
radial and azimuthal components which depend on the spherical coordinates
$r$ and $\theta$. The radial component $v_{r}\left(r\right)$ is
assumed to be in the classical beta-velocity law \citep[c.f.][]{castor:1979},
and the azimuthal component $v_{\phi}$ is assumed to be a constant
fraction ($\gamma$) of the Keplerian velocity for a given distance
($w=\sqrt{x^{2}+y^{2}}$) from the symmetry axis ($z$-axis) , i.e.:

\begin{eqnarray}
\boldsymbol{v}_{\mathrm{wind}} & = & v_{r}\hat{\boldsymbol{r}}+v_{\phi}\hat{\boldsymbol{\bphi}}\label{eq:wind_velocity_def}\end{eqnarray}
where \begin{equation}
v_{r}\left(r\right)=v_{r0}+\left(v_{\infty}-v_{r0}\right)\left(1-\frac{R_{\mathrm{mo}}}{r}\right)^{\beta}\,,\label{eq:wind_velocity_radial_part}\end{equation}
\begin{equation}
v_{\phi}\left(w\right)=\gamma\,\left(\frac{GM_{*}}{w}\right)^{1/2}\,,\label{eq:wind_velocity_azimuth_part}\end{equation}
Note that the base of the wind starts at $r=R_{\mathrm{mo}}$. This
is chosen so mainly for computational simplicity i.e. to avoid the
overlap between the wind and the magnetosphere; however, see \citet{matt:2005}
for the possibility that wind originates from the stellar surface
(restricted near the polar caps). The range of the polar angle for
the wind is restricted to $\theta>\left|\theta_{\mathrm{disc}}\right|$
where $\theta_{\mathrm{disc}}$ is the opening angle of the accretion
disc, to avoid the overlap. $v_{r0}$ is the small radial velocity
at the base of the wind ($r=R_{\mathrm{mo}}$). Normally, $v_{r0}=10\,\mathrm{km\, s^{-1}}$
which is approximately equal to the thermal velocity of hydrogen with
$T=7500\,\mathrm{K}$. Following \citet{appenzeller:2005}, $\gamma=0.05$
is adopted. The dependency of $v_{r}$ in polar direction ($\theta=0$)
on the values of wind acceleration parameter $\beta$ is shown in
Figure~\ref{fig:wind_vr_rho}. All other parameters describing the
wind are fixed as the standard values given in Table~\ref{tab:std_parameters}.

@@

\caption{The dependency of the bipolar wind velocity and density
structures on the wind acceleration parameter $\beta$. The radial
component of the wind velocity
(equation~\ref{eq:wind_velocity_radial_part}) in polar direction,
$\theta = 0$ as a function of radius is shown on left. The wind
density (equation~\ref{eq:wind_density_final} along polar direction as
a function of radius is shown on right. For all $\beta$ values, the
density is normalised to the density at the base of the wind in polar
direction, i.e. $\rho_0 = \rho\left(R_\mathrm{mo},0\right)$.  The
smaller the value of $\beta$, the faster the acceleration of the
wind. For the radius ($r/R_{\mathrm{mo}} < 10$, both velocity and
density are sensitive to the value of $\beta$. For all $\beta$ values,
the radial velocity of the wind reaches the terminal velocity ($\sim
210\,\kmps$ by $r/R_{\mathrm{mo}} \sim 1000$. The initial velocity
$V_{0}=10\,\kmps$, which approximately corresponds to the thermal
velocity of a hydrogen atom at 7500~K, is used for all $v_{r}$
plots. Beyond $r/R_{\mathrm{mo}}= 100$, little difference is seen in
the polar density with different $\beta$ values.}

@@

To calculate the dust scattering and absorption cross section as a
function of wavelength, the optical constants of \citet{draine:1984}
for amorphous carbon grains and \citet{hanner:1988} for silicate
grains are used. The model uses the {}``large grain'' dust model
of \citet{wood:2002} in which the dust grain size distribution is
described by the following function:\begin{equation}
n\left(a\right)da=\left(C_{\mathrm{C}}+C_{\mathrm{Si}}\right)\, a^{-p}\exp\left[-\left(\frac{a}{a_{c}}\right)^{q}\right]da\label{eq:grain-dist-function}\end{equation}
 where $a$ is the grain size restricted between $a_{\mathrm{min}}$
and $a_{\mathrm{max}}$, and $C_{\mathrm{C}}$ and $C_{\mathrm{Si}}$
are the terms set by requiring the grains to completely deplete a
solar abundance carbon and silicon. The parameters adopted in our
model are: $C_{\mathrm{C}}=1.32\times10^{-17}$, $C_{\mathrm{Si}}=1.05\times10^{-17}$,
$p=3.0$, $q=0.6$, $a_{\mathrm{\mathrm{min}}}=0.1\,\mathrm{\mathrm{\mathrm{\mu m}}}$,
$a_{\mathrm{\mathrm{max}}}=1000\,\mathrm{\mathrm{\mathrm{\mu m}}}$,
and $a_{c}=50\,\mathrm{\mathrm{\mathrm{\mu m}}}$. This corresponds
to Model~1 of the dust model used by \citet{wood:2002}. See also
their Fig.~3 The relative number of each grain is assumed to be that
of solar abundance, C/H$\sim 3.5\times10^{-4}$ \citep{anders:1989}
and Si/H$\sim 3.6\times10^{-5}$ \citep{grevesse:1993} which are similar
to values found in the ISM model of \citet{mathis:1977} and \citet*{kim:1994}.
Similar abundances were used in the circumstellar disc models of \citet{cotera:2001}.

@@

\caption{Basic Model Configuration of the disc-wind, magnetosphere
hybrid model. The system consist of four components: (1)~the continuum
source located at the origin $\left( o \right)$ of the cartesian
coordinates $\left(x,y,z\right)$ -- the y-axis are into the paper,
(2)~magnetospherical accretion flow, (3)~collimated wind/jet
outflow, and (4)~geometrically thin (but opaque) accretion disc. The
disc wind originates from the disc surface between $w_{i}=R_{wi}$ and
$w_{i}=R_{wo}$.  The wind source points ($S$), from which the stream
lines diverges, are placed at distance $d$ above and below the
star. By changing the value of $d$, the degree of the wind collimation
is controlled. The density distribution is symmetric around the
z-axis. }
